Comparative safety and effectiveness of Pfizer BA.4-5 versus Sanofi during the spring 2023 COVID-19 booster vaccination programme in England: a matched cohort study in OpenSAFELY-TPP

### 2024-02-27 write-up/manuscript/manuscript.qmd
manuscript edits

TBD: Colm D Andrews^1^, Edward PK Parker^2^, Em Prestige^2^, Venexia Walker^3,4^, Tom Palmer^3,4^, Helen J Curtis^1^, Alex Walker^1^, Amir Mehrkar^1^, Brian MacKenna^1^, Sebastian CJ Bacon^1^, Ben Goldacre^1^, Miguel A Hernán^5,6^\*, Jonathan AC Sterne^3,7,8^\*, William J Hulme^1^, and the OpenSAFELY collaborative.

@@

*Introduction* The spring 2023 COVID-19 booster vaccination programme in England used both Pfizer BA.4-5 and Sanofi vaccines. All people aged 75 years or over and the clinically vulnerable were eligible to receive a booster dose. Direct comparisons of effectiveness against severe COVID-19 events of these two vaccines for boosting have not been made in trials or observational data.

*Methods* With the approval of NHS England, we used the OpenSAFELY-TPP database to compare effectiveness of the Pfizer BA.4-5 and Sanofi vaccines during the spring 2023 booster programme, between 1 April and 30 June 2023. We investigated two cohorts separately: those aged 75 or over (75+); those aged 50 or over and clinically vulnerable (CV). In each cohort, vaccine recipients were matched on date of vaccination, COVID-19 vaccine history, age, and other characteristics. Effectiveness outcomes were COVID-19 hospital admission, COVID-19 critical care admission, and COVID-19 death up to 16 weeks after vaccination. Safety outcomes were Pericarditis and Myocarditis up to 4 weeks after vaccination. We report the cumulative incidence of each outcome, and compare safety and effectiveness using risk differences (RD), relative risks (RR), and incidence rate ratios (IRRs).

*Results* `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5") %>% pull(n) %>% sum())` people were 1-1 matched in the CV cohort, and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5") %>% pull(n) %>% sum())` in the 75+ cohort, contributing a total `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up respectively. The incidence of COVID-19 hospital admission was higher for Sanofi than for Pfizer BA.4-5. In the CV cohort, 16-week risks per `r perN_format` people were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi, with an IRR of `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)`. In the 75+ cohort, these were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi, with an IRR of `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI)`. These finding were similar across all pre-specified subgroups. More severe COVID-19 related outcomes (critical care admission and death), and safety outcomes at 4 weeks, were rare in both vaccines so we could not reliably estimate superiority of either vaccine.

*Conclusion* This observational study comparing effectiveness of Pfizer BA.4-5 and Sanofi vaccine during the spring 2023 programme in England in the two main eligible cohorts -- people aged 75 and over and in clinically vulnerable people -- found some evidence of superior effectiveness against COVID-19 hospital admission for Pfizer BA.4-5 compared with Sanofi within 16 weeks after vaccination. Differences were largely observed during weeks 2 to 4 after vaccination, suggesting the Pfizer BA.4-5 offers earlier protection via more rapid immunogenicity, as shown in RCTs.

@@

The spring 2023 COVID-19 booster vaccination programme in England offered COVID-19 vaccination to those aged 75 years or over, resident in a care home, and those who are immunosuppressed, based on guidance from the Joint Committee for Vaccination and Immunisation (JCVI).[@NHSEnglandNHSc; @JCVIStatementCOVID19; @JCVIStatementSpring] Both the Pfizer BA.4-5 vaccine (authorised November 2022 [@SecondPfizerBioNTech]) and the Sanofi vaccine (authorised December 2022 [@SanofiPasteurCOVID19]) were used during this programme. This allows direct comparisons of their safety and effectiveness against severe COVID-19 disease, something not investigated in randomised trials.

@@

All data were linked, stored and analysed securely using the OpenSAFELY platform: <https://opensafely.org/>, as part of the NHS England OpenSAFELY COVID-19 service. Data include pseudonymised data such as coded diagnoses, medications and physiological parameters. No free text data are included. All code is shared openly for review and re-use under MIT open license <https://github.com/opensafely/comparative-booster-spring2023>. Detailed pseudonymised patient data is potentially re-identifiable and therefore not shared. Primary care records managed by the GP software provider, TPP were linked, using NHS numbers, to emergency department attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS). COVID-19 vaccination history and health and social care worker status is available in the GP record directly via the National Immunisation Management System (NIMS).

@@

We considered two cohorts: all adults aged 75 years or over ("75+"), and all adults aged 50 or over who were considered clinically vulnerable ("CV"). These groups are not mutually exclusive. We only considered those aged 50 years or over because the Sanofi vaccine was rarely offered to younger people. Clinical vulnerability encompasses various chronic or immunosuppressive conditions and medications as set out in the Green Book chapter 14a @COVID19GreenBook, and defined in supplementary materials.

@@

Age was calculated as at 30 June 2023, to allow for a degree of operational flexibility in who was offered vaccination during the programme. All other characteristics were calculated as on the day before vaccination. In the analysis cohorts there are no missing values for any characteristics as people with missing values for sex, IMD or STP were excluded, and remaining characteristics were defined by the presence or absence of clinical codes or events in the health record. The supplementary materials provide more information on how these characteristics were defined.

@@

We considered two safety outcomes. Pericarditis within 28 days of vaccination was identified using SUS emergency department records with pericarditis listed as the discharge diagnosis (SNOMED), and using ICD-10 coded hospital admission and death records. Myocarditis was identified similarly. Details of the codelists used to define these outcomes are available in supplementary materials.

@@

Baseline characteristics of each vaccine group were tabulated, with between-group balance examined using standardised mean differences. We estimated the cumulative incidence of each outcome in each vaccine group using the Kaplan-Meier (KM) estimator. We estimated cumulative risk differences and risk ratios (RRs) comparing the vaccine groups for each outcome. We also estimated Incidence Rate Ratios (IRRs) over the full follow-up duration, and within period-specific intervals defined by splits on weeks 1, 2, 4, 8, 12, and 16.

@@

### Software, code, and reproducibility

Data management and analyses were conducted using OpenSAFELY tools, Python version 3.8.10 and R version 4.0.2. Code for data management and analysis, as well as codelists, are archived online <https://github.com/opensafely/comparative-booster-spring2023>. Codelists are available at <https://www.opencodelists.org/>. The supplementary materials provide further details of codelists and data sources used for all variables in the study. Detailed pseudonymised patient data is potentially re-identifiable and therefore not shared. \newpage

## Results

@@

Sanofi was more likely to be administered in April 2023, and Pfizer BA.4-5 more likely in May and June 2023. Prior to matching, Sanofi recipients were on average older, more likely to be recorded as resident in a care or nursing home, and less likely to have received fewer than 5 prior COVID-19 vaccine doses compared with Pfizer BA.4-5 recipients. Sanofi recipients were less likely that Pfizer BA.4-5 recipients to have previously received the Pfizer/BA.1 vaccine during an earlier booster programme and more likely to have received Moderna/Omicron. Sex, deprivation levels, and ethnicity were similar in both groups prior to matching (supplementary Tables S1a and S1b).

In the CV cohort, `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(n))` people in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`) were matched, representing `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c5", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 1a, Figure S1a). In the 75+ cohort, `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(n))` were matched in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`), representing `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c5", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 1b, Figure S1b).

@@

The 16-week risks (cumulative incidence) per `r perN_format` people of COVID-19 hospital admissions were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(rdCI_95)`) (Table 2; Figure 3). The corresponding 16-week risks per `r perN_format` people for COVID-19 critical care admission were `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(rdCI)`). The 16-week risks per `r perN_format` people of COVID-19 death were `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(rdCI)`). For non-COVID-19 death, these were `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(rdCI)`).

IRRs (where less than one favours Sanofi) in the 16-week period after vaccination were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admissions, `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 3).

@@

@@

In the CV cohort, there were `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_1)` pericarditis events within 4 weeks after vaccination for Pfizer BA.4-5 and Sanofi recipients respectively. This corresponded to a cumulative incidence per `r perN_format` people of `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(rdCI_95)`). For myocarditis, there were `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` events for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events for Sanofi, (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`)

@@

In the 75+ cohort, `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_1)` pericarditis events within 4 weeks after vaccination for Pfizer BA.4-5 and Sanofi recipients respectively, with a cumulative incidence per `r perN_format` people of `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(rdCI_95)`). For myocarditis, there were `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` for Pfizer BA.4/5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events for Sanofi, (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`).

@@

Findings in the main analyses were similar when examining subgroups, those estimated with reduced precision. Supplementary materials provide cumulative incidence plots by subgroup (Figures S3), the corresponding cumulative risk differences, cumulative risk ratios, and period-specific IRRs (Figures S4), and subgroup-specific heterogeneity tests for risk differences and risk ratios (Table S2).

@@

This observational study of COVID-19 vaccine recipients during the spring 2023 booster programme in England is, to our knowledge, the first to directly compare the Pfizer BA.4-5 and Sanofi vaccines for safety and effectiveness against severe COVID-19 outcomes. We estimated that, in both CV (N=`r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) and 75+ (N=`r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) cohorts, COVID-19 related hospital admissions were less frequent in those receiving the Pfizer BA.4-5 vaccine compared with Sanofi during the first 16 weeks after vaccination. For the more severe COVID-19 outcomes of critical care admission and death, there were very few events and insufficient power to identify differences between vaccines.

@@

The OpenSAFELY-TPP database covers around 45% of the English population and contains rich clinical information which enabled us to closely match Pfizer BA.4-5 and Sanofi recipients to control for potential confounding and compare comparative safety and effectiveness.

@@

<!-- The MHRA authorisation for Sanofi was based on clinical data from about 800 individuals previously immunised with an mRNA or viral-vectored vaccine \[@SanofiPasteurCOVID19\] and authorisation for Pfizer BA.4/BA.5 was based on a placebo-controlled efficacy trial, finding efficacy of 86 to 100% against laboratory-confirmed COVID-19. [@zouNeutralizationBABA2023] -->

Out-of-trial evidence of the effectiveness of Sanofi and Pfizer BA.4-5 vaccines against hospitalisation amongst adults aged 75 years and older in England from a test-negative case control study involving 14,174 eligible tests found similar effectiveness of both Pfizer BA.4-5 and Sanofi of 30-50%.[@kirsebomEffectivenessAdjuvantedSanofi2023] We are aware of no other planned or published randomised trials making direct endpoint comparisons of boosting with Pfizer BA.4-5 and Sanofi.

### Conclusions

During the spring 2023 COVID-19 booster programme in England, COVID-19-related hospital admissions were slightly more frequent following receipt of the Sanofi vaccine than the Pfizer BA.4-5 vaccine after matching on a number of important potential confounders. These findings were broadly consistent across subgroups, though estimated with less precision.

>>>>

TBD: Colm D Andrews^1^, Em Prestige^2^, Edward PK Parker^2^, Venexia Walker^3,4^, Tom Palmer^3,4^, Helen J Curtis^1^, Alex Walker^1^, Amir Mehrkar^1^, Brian MacKenna^1^, Sebastian CJ Bacon^1^, Ben Goldacre^1^, Miguel A Hernán^5,6^\*, Jonathan AC Sterne^3,7,8^\*, William J Hulme^1^, and the OpenSAFELY collaborative.

@@

*Introduction* The spring 2023 COVID-19 booster vaccination programme in England used both Pfizer BA.4-5 and Sanofi vaccines. All people aged 75 years or over and the clinically vulnerable were eligible to receive a booster dose. Direct comparisons of these two vaccines in boosting protection against severe COVID-19 events have not been made in trials or observational data.

*Methods* With the approval of NHS England, we used the OpenSAFELY-TPP database to compare effectiveness of the Pfizer BA.4-5 and Sanofi vaccines during the spring 2023 booster programme, between 1 April and 30 June 2023. We investigated two cohorts separately: those aged 75 or over (75+); and those aged 50 or over and clinically vulnerable (CV). In each cohort, vaccine recipients were matched on date of vaccination, COVID-19 vaccine history, age, and other characteristics. Effectiveness outcomes were COVID-19 hospital admission, COVID-19 critical care admission, and COVID-19 death up to 16 weeks after vaccination. Safety outcomes were pericarditis and myocarditis up to 4 weeks after vaccination. We report the cumulative incidence of each outcome, and compare safety and effectiveness using risk differences (RD), relative risks (RR), and incidence rate ratios (IRRs).

*Results* `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5") %>% pull(n) %>% sum())` people were 1-1 matched in the CV cohort, and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5") %>% pull(n) %>% sum())` in the 75+ cohort, contributing a total `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up, respectively. The incidence of COVID-19 hospital admission was higher for Sanofi than for Pfizer BA.4-5. In the CV cohort, 16-week risks per `r perN_format` people were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi, with an IRR of `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)`. In the 75+ cohort, these were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi, with an IRR of `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI)`. These finding were similar across all pre-specified subgroups. More severe COVID-19 related outcomes (critical care admission and death), and safety outcomes at 4 weeks, were rare in both vaccines so we could not reliably compare effectiveness of the two vaccines.
*Conclusion* This observational study comparing effectiveness of Pfizer BA.4-5 and Sanofi vaccine during the spring 2023 programme in England in the two main eligible cohorts -- people aged 75 and over and in clinically vulnerable people -- found some evidence of superior effectiveness against COVID-19 hospital admission for Pfizer BA.4-5 compared with Sanofi within 16 weeks after vaccination. Differences were largely observed during weeks 2 to 4 after vaccination, suggesting the Pfizer BA.4-5 offers earlier protection via more rapid immunogenicity, as shown in RCTs.

@@

The spring 2023 COVID-19 booster vaccination programme in England offered COVID-19 vaccination to those aged 75 years or over, resident in a care home, and those who are immunosuppressed, based on guidance from the Joint Committee for Vaccination and Immunisation (JCVI).[@NHSEnglandNHSc; @JCVIStatementCOVID19; @JCVIStatementSpring] Both the Pfizer BA.4-5 vaccine (authorised November 2022 [@SecondPfizerBioNTech]) and the Sanofi vaccine (authorised December 2022 [@SanofiPasteurCOVID19]) were used during this programme. This allows direct comparisons of their safety and effectiveness against severe COVID-19 disease - something not investigated in randomised trials.

@@

All data were linked, stored and analysed securely using the OpenSAFELY platform: <https://opensafely.org/>, as part of the NHS England OpenSAFELY COVID-19 service. Data include pseudonymised data such as coded diagnoses, medications and physiological parameters. No free text data are included. All code is shared openly for review and re-use under MIT open license (<https://github.com/opensafely/comparative-booster-spring2023>). Detailed pseudonymised patient data is potentially re-identifiable and therefore not shared. Primary care records managed by the GP software provider, TPP were linked, using NHS numbers, to emergency department attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS). COVID-19 vaccination history and health and social care worker status is available in the GP record directly via the National Immunisation Management System (NIMS).

@@

We considered two cohorts: all adults aged 75 years or over ("75+"), and all adults aged 50 or over who were considered clinically vulnerable ("CV"). These groups are not mutually exclusive. We only considered those aged 50 years or over because the Sanofi vaccine was rarely offered to younger people. Clinical vulnerability encompasses various chronic or immunosuppressive conditions and medications as set out in the Green Book chapter 14a @COVID19GreenBook, and defined in the supplementary materials.

@@

Age was calculated as of 30 June 2023, to allow for a degree of operational flexibility in who was offered vaccination during the programme. All other characteristics were calculated as of the day before vaccination. In the analysis cohorts there are no missing values for any characteristics as people with missing values for sex, IMD or STP were excluded, and remaining characteristics were defined by the presence or absence of clinical codes or events in the health record. The supplementary materials provide more information on how these characteristics were defined.

@@

We considered two safety outcomes. Pericarditis within 28 days of vaccination was identified using SUS emergency department records with pericarditis listed as the discharge diagnosis (SNOMED), and using ICD-10 coded hospital admission and death records. Myocarditis was identified similarly. Details of the codelists used to define these outcomes are available in the supplementary materials.

@@

Baseline characteristics of each vaccine group were tabulated, with between-group balance examined using standardised mean differences. We estimated the cumulative incidence of each outcome in each vaccine group using the Kaplan-Meier (KM) estimator. We estimated cumulative risk differences and risk ratios (RRs) comparing the vaccine groups for each outcome. We also estimated Incidence Rate Ratios (IRRs) for Sanofi versus Pfizer BA.4-5 recipients over the full follow-up duration, and within period-specific intervals defined by splits on weeks 1, 2, 4, 8, 12, and 16.

@@

### Software, code, and reproducibility
Data management and analyses were conducted using OpenSAFELY tools, Python version 3.8.10 and R version 4.0.2. Code for data management and analysis, as well as codelists, are archived online (<https://github.com/opensafely/comparative-booster-spring2023>). Codelists are available at <https://www.opencodelists.org/>. The supplementary materials provide further details of codelists and data sources used for all variables in the study. Detailed pseudonymised patient data is potentially re-identifiable and therefore not shared. \newpage
## Results

@@

Sanofi was more likely to be administered in April 2023, and Pfizer BA.4-5 more likely in May and June 2023. Prior to matching, Sanofi recipients were on average older, more likely to be recorded as resident in a care or nursing home, and less likely to have received fewer than 5 prior COVID-19 vaccine doses compared with Pfizer BA.4-5 recipients, in both cohorts. Sanofi recipients were less likely that Pfizer BA.4-5 recipients to have previously received the Pfizer/BA.1 vaccine during an earlier booster programme and more likely to have received Moderna/Omicron. Sex, deprivation levels, and ethnicity were similar in both groups prior to matching (supplementary Tables S1a and S1b).
In the CV cohort, `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(n))` people in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`) were matched, representing `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c5", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 1a, Figure S1a). In the 75+ cohort, `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(n))` were matched in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`), representing `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c5", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 1b, Figure S1b).

@@

The 16-week risks (cumulative incidence) per `r perN_format` people of COVID-19 hospital admissions were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(rdCI_95)`) (Table 2; Figure 3). The corresponding 16-week risks per `r perN_format` people for COVID-19 critical care admission were `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(rdCI)` where values greater than zero favour Pfizer). The 16-week risks per `r perN_format` people of COVID-19 death were `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(rdCI)`). For non-COVID-19 death, these were `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(rdCI)`).

IRRs (where values greater than one favour Pfizer) in the 16-week period after vaccination were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admissions, `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 4).

@@

```{r break4}

@@

In the CV cohort, there were `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_1)` pericarditis events within 4 weeks after vaccination for Pfizer BA.4-5 and Sanofi recipients, respectively. This corresponded to a cumulative incidence per `r perN_format` people of `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(rdCI_95)`). For myocarditis, there were `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` events for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`)

@@

In the 75+ cohort, `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_1)` pericarditis events within 4 weeks after vaccination for Pfizer BA.4-5 and Sanofi recipients, respectively, with a cumulative incidence per `r perN_format` people of `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(rdCI_95)`). For myocarditis, there were `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` events for Pfizer BA.4/5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`).

@@

Findings in the main analyses were similar when examining subgroups, although estimated with reduced precision. The Supplementary materials provide cumulative incidence plots by subgroup (Figures S3), the corresponding cumulative risk differences, cumulative risk ratios, and period-specific IRRs (Figures S4), and subgroup-specific heterogeneity tests for risk differences and risk ratios (Table S2).

@@

This observational study of COVID-19 vaccine recipients during the spring 2023 booster programme in England is, to our knowledge, the first to directly compare the Pfizer BA.4-5 and Sanofi vaccines for safety and effectiveness against severe COVID-19 outcomes. We estimated that, in both CV (N=`r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) and 75+ (N=`r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) cohorts, COVID-19 related hospital admissions were less frequent in those receiving the Pfizer BA.4-5 vaccine compared with Sanofi during the first 16 weeks after vaccination. For the more severe COVID-19 outcomes of critical care admission and death, there were very few events resultng in reduced power to identify differences between vaccines.

@@

The OpenSAFELY-TPP database covers around 45% of the English population and contains rich clinical information which enabled us to closely match Pfizer BA.4-5 and Sanofi recipients to control for potential confounding and compare safety and effectiveness.

@@

<!-- The MHRA authorisation for Sanofi was based on clinical data from about 800 individuals previously immunised with an mRNA or viral-vectored vaccine \[@SanofiPasteurCOVID19\] and authorisation for Pfizer BA.4/BA.5 was based on a placebo-controlled efficacy trial, finding efficacy of 86 to 100% against laboratory-confirmed COVID-19. [@zouNeutralizationBABA2023] -->
Out-of-trial evidence of the effectiveness of Sanofi and Pfizer BA.4-5 vaccines against hospitalisation amongst adults aged 75 years and older in England from a test-negative case control study involving 14,174 eligible tests found similar effectiveness of both Pfizer BA.4-5 and Sanofi of 30-50%.[@kirsebomEffectivenessAdjuvantedSanofi2023] We are aware of no other planned or published randomised trials making direct endpoint comparisons of boosting with Pfizer BA.4-5 and Sanofi. 
### Conclusions

During the spring 2023 COVID-19 booster programme in England, COVID-19-related hospital admissions were slightly more frequent following receipt of the Sanofi vaccine than the Pfizer BA.4-5 vaccine after matching on a number of important potential confounders. These finding were broadly consistent across subgroups, though estimated with less precision.
